Self Regulation Ontology Retest Data: Initial Exploration

### 2017-02-26 SRO_Retest_Analyses.Rmd
cont working on retest exploration

What proportion is not normal?  
A large proportion of the variables are not normal for data from both sources. Larger proportions of survey data are not normal compared to task data. 
```{r}
dist_data %>%
  group_by(task) %>%
  summarise(mean_retest_not_normal  = mean(retest_not_normal),
            mean_test_not_normal = mean(test_not_normal))
```

Are the same vars that are not normally distributed in the retest data not normal in test data as well?  
List of variables that have different distributions in the two time points.
A majority of them are drift diffusion parameters (but note that these vars are overrepresented to begin with - more on them later).
```{r}
dist_data %>%
  filter(retest_not_normal != test_not_normal)
```

@@

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(positive_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.logTr')
  cat(rep('*', 40))
  cat(paste0('Dropping ', length(dropped_vars) ,' positively skewed data that could not be transformed successfully:'))
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  data = cbind(data, successful_transforms)

@@

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(negative_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.ReflogTr')
  cat(rep('*', 40))
  cat(paste0('Dropping ', length(dropped_vars) ,' negatively skewed data that could not be transformed successfully:'))
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  data = cbind(data, successful_transforms)

@@

```{r}
clean_all_columns <- unique(c(names(clean_retest_data), names(clean_retest_subs_test_data)))

@@

for(i in 1:length(clean_matching_dv_columns)){
  clean_cor_df[clean_matching_dv_columns[i], 'pearson'] <- cor(clean_retest_data[,clean_matching_dv_columns[i]], clean_retest_subs_test_data[,clean_matching_dv_columns[i]], method = 'pearson', use = "pairwise.complete.obs")
  clean_cor_df[clean_matching_dv_columns[i], 'spearman'] <- cor(clean_retest_data[,clean_matching_dv_columns[i]],clean_retest_subs_test_data[,clean_matching_dv_columns[i]], method = 'spearman', use = "pairwise.complete.obs")
}
```

```{r}
clean_cor_df %>%
  gather(key, value) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~key)
```

```{r}
clean_cor_df$dv = row.names(clean_cor_df)
row.names(clean_cor_df) = seq(1:nrow(clean_cor_df))
clean_cor_df$task = 'task'
clean_cor_df[grep('survey', clean_cor_df$dv), 'task'] = 'survey'
```

```{r}
clean_cor_df %>%
  gather(key, value, -dv, -task) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(key~task)
```

This cleaning led to a slight change in the mean reliabilities but a lot of dropped variables.

@@

Drift diffusion variables: raw vars vs. EZ vs. hddm

>>>>

**What proportion is not normal?**  
A large proportion of the variables are not normal for data from both sources. Larger proportions of survey data are not normal compared to task data. 
```{r}
dist_data %>%
  group_by(task) %>%
  summarise(mean_retest_not_normal  = mean(retest_not_normal),
            mean_test_not_normal = mean(test_not_normal))
```

**Are the same vars that are not normally distributed in the retest data not normal in test data as well?**  
List of variables that have different distributions in the two time points.
A majority of them are drift diffusion parameters (but note that these vars are overrepresented to begin with - more on them later).
```{r}
dist_data %>%
  filter(retest_not_normal != test_not_normal)
```

@@

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(positive_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.logTr')
  cat(rep('*', 40))
  cat('\n')
  cat(paste0('Dropping ', length(dropped_vars) ,' positively skewed data that could not be transformed successfully:'))
  cat('\n')
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  cat('\n')
  data = cbind(data, successful_transforms)

@@

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(negative_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.ReflogTr')
  cat(rep('*', 40))
  cat('\n')
  cat(paste0('Dropping ', length(dropped_vars) ,' negatively skewed data that could not be transformed successfully:'))
  cat('\n')
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  cat('\n')
  data = cbind(data, successful_transforms)

@@

Create the same list of correlations as before but this time on a smaller set of cleaned variables.
```{r}
clean_all_columns <- unique(c(names(clean_retest_data), names(clean_retest_subs_test_data)))

@@

for(i in 1:length(clean_matching_dv_columns)){
  clean_cor_df[clean_matching_dv_columns[i], 'pearson'] <- cor(clean_retest_data[,clean_matching_dv_columns[i]], clean_retest_subs_test_data[,clean_matching_dv_columns[i]], method = 'pearson', use = "pairwise.complete.obs")
  clean_cor_df[clean_matching_dv_columns[i], 'spearman'] <- cor(clean_retest_data[,clean_matching_dv_columns[i]],clean_retest_subs_test_data[,clean_matching_dv_columns[i]], method = 'spearman', use = "pairwise.complete.obs")
}

clean_cor_df$dv = row.names(clean_cor_df)
row.names(clean_cor_df) = seq(1:nrow(clean_cor_df))
clean_cor_df$task = 'task'
clean_cor_df[grep('survey', clean_cor_df$dv), 'task'] = 'survey'
```

Let's look at the distributions again: The distributions don't look terribly different but there are noticably fewer variables post cleaning. 
```{r warning=FALSE, message=FALSE}
cor_df %>%
  left_join(clean_cor_df, by='dv') %>%
  select(-task.y)%>%
  rename(pearson_preClean = pearson.x, spearman_preClean = spearman.x,
         pearson_postClean = pearson.y, spearman_postClean = spearman.y,
         task = task.x) %>%
  gather(key, value, -dv, -task) %>%
  separate(key, c('cor_type', 'clean_type'),sep = "_") %>%
  ggplot(aes(x=value, fill=factor(clean_type, levels = c('preClean', 'postClean'),
                                  labels = c('Before', 'After'))))+
  geom_histogram(position='identity', alpha = 0.5)+
  facet_wrap(cor_type~task)+
  theme_bw()+
  theme(legend.title = element_blank())
```
This cleaning led to a slight change in the mean reliabilities but a lot of dropped variables.

@@

**What changed most with cleaning?**    
Difference in correlation pre and post correlation for the variables that survived cleaning.  
Things that changed most seem to be RTs and parameter estimates.

```{r}
clean_cor_df %>%
  left_join(cor_df, by='dv') %>%
  mutate(pearson_diff = abs(pearson.y - pearson.x),
         spearman_diff = abs(spearman.y - spearman.x)) %>%
  select(dv, pearson_diff, spearman_diff) %>%
  arrange(-pearson_diff, -spearman_diff)
```

# Different variable types

##

## Drift diffusion variables: raw vars vs. EZ vs. hddm
